Learning constraints in tabular data

### 2016-05-27 Paper/ECAI/tabular.tex
Further comments Angelika

\begin{example}
Consider the (headerless) table~$T_1$ in Figure~\ref{fig:main_example}, its rows are not type consistent (i.e. they contain both numeric and textual data).
However, the table can be partitioned into five column-oriented blocks $B_1, B_2, B_3, B_4, B_5$, as shown in the figure ($B_1 = \range{T_1}{\rangeall}{1}$, $B_2 = \range{T_1}{\rangeall}{2}$, $B_3 = \range{T_1}{\rangeall}{\rangeto{3}{8}}$, \dots).
% {\small
% B_1 = \range{T_1}{\rangeall}{1},
% B_2 = \range{T_1}{\rangeall}{2},
% B_3 = \range{T_1}{\rangeall}{\rangeto{3}{8}},
% B_4 = \range{T_1}{\rangeall}{9},
% B_5 = \range{T_1}{\rangeall}{\rangeto{10}{11}}
\end{example}

@@

Instead, we divide this step into two parts: in the first part, we will not reason over individual (sub)blocks and their data, but rather over the properties of the maximal superblocks. % and what combination of superblocks may lead to a valid argument assignment.\\
Consider Table 1 and 2 in Figure~\ref{fig:main_example} and the $SUM_{col}$ constraint template. Instead of considering all possible subblocks of T1 and T2, we reason over the properties of the superblocks first, $B2$ and $B4$ are not numeric so can be immediately discarded. Looking at the sizes, $B1$ contains only one column and $B5$ 2, however, the rows in $T2$ have length 4 so we only need to consider superblocks of size at least 4, which only $B3$ satisfies.

In the second part, we start from an atom $c(B_1, \ldots ,B_n)$ with maximal superblocks $B_i$, and generate and test all possible (sub)block assignments to arguments using the properties and actual data of the blocks. As we only have to consider the blocks $B'_i \sqsubseteq B_i$ contained in each superblock, this is typically much easier. For the above example, each of the vectors of $B6$ will be considered as candidate for the left-hand side, and one can enumerate all subblocks for the right-hand side and verify the signature (e.g. at least size $4$) and test the definition for each of the rows. %In practice, this can be optimized further.
\\\\
We now describe in more detail how the headerless tables are extracted and how the maximal superblocks are generated from them (step 1 and 2, Section 4.1), how the candidate superassignments are generated (step 3a, Section 4.2) and how the actual assignments are extracted from that (step 3b, Section 4.3). We then describe a number of optimizations that we perform to speed up step 3 and 4 (Section 4.4).

@@

Table~\ref{tbl:translation} shows how the conversion from requirements of the signature to CSP constraints happens:
Requirements on the lengths, orientations and tables of blocks can be directly enforced since they are invariant under block containment($\sqsubseteq$).
Typing requirements need to be relaxed to check only for base types (numeric and textual).
Minimum sizes can be directly enforced, but exact size requirements are relaxed to minimum sizes, since subgroups might contains fewer vectors.
Finally, restrictions on the number of rows or columns behave as length or size constraints based on the orientation of the group they are applied to.
Subgroups of row-oriented (column-oriented) group will always have the same number of columns (rows), however, the number of rows (columns) might decrease.

The $\generategroups(\textit{s,\blocks})$ method will use these conversion rules to construct a CSP and query a solver for all satisfying solutions.
These solutions correspond to the valid superblock assignments for constraint template~$s$.

@@

In this step, we could also formulate the problem as a CSP. %However, the search space is much smaller, e.g. $O(nm^2)$ with $n$ the number of arguments of the template and $m$ the size of the largest block.
However, few CSP solvers support floating point numbers, which are prevalent in spreadsheets. Furthermore, in a CSP approach we would have to \textit{ground} out the definition for each of the corresponding elements in the vectors. This is inefficient, as the signature of the definition may already not be satisfied, or the first element in the data.

@@

Using the dependency graph~(\dependencies) we can reorder the templates such that a template occurs after any template it depends on.
Concretely, the input templates~$\temps$ (see Algorithm~\ref{algo:tcl}) are reordered to agree with the partial ordering imposed by~\dependencies.

The dependency graph represents a partial ordering and is used produce a linear ordering of the given templates such that more general templates occur before templates that depend on them (e.g. \textit{ALLDIFFERENT} before \textit{FOREIGNKEY}).
Considering Algorithm~\ref{algo:tcl}, this step can be seen as preprocessing the input~\temps.
Learning in this order allows templates to use constraints they depend on as these will have already been found.
This is exploited in the \generategroups step to speedup the algorithm by reducing the search space.
Let us illustrate this step with an example.

@@

\subsubsection{Redundancy}
We consider two types of redundancy that we aim to eliminate during the search.
As noted in Section~\ref{sec:form:dependencies}, for some constraint templates there are \textit{symmetries} over the arguments that lead to trivially equivalent solutions, for example, $\ecprod{B_R'}{B_1'}{B_2'} \Leftrightarrow \ecprod{B_R'}{B_2'}{B_1'}$.
Such duplicates are avoided by defining a \textit{canonical} form for these constraints.
In practice, we define an arbitrary but fixed block ordering and require those blocks that are interchangeable to adhere to this ordering.
Moreover, \sname currently contains only product (\ecprod{B_R'}{B_1'}{B_2'}) and not division (\eccalc{B_1'}{B_R' / B_2'}) as well as difference (\ecdiff{B_r}{B_1}{B_2}) and not sum (\eccalc{B_1}{B_r + B_2}).
These choices are based on the prevalence of product and difference in financial spreadsheet data, however, division and sum could easily be added to, as can other constraints.

@@

\subsubsection{Constraints}
The constraints that are currently supported in our system are shown in Table~\ref{table:constraints}.
Their inclusion was based on the analysis of the most popular Excel constraints as well as the spreadsheets we encountered.
All these constraints are available in popular spreadsheet software, except for the structural constraints (\textit{ALLDIFFERENT}, \textit{PERMUTATION} and \textit{FOREIGNKEY}).
These constraints added so they can be used in the signature of other constraint templates; they are also popular in constraint satisfaction~\cite{modelseeker}.

@@

\section{Evaluation}\label{sec:evaluation}
\sergey{we need to add a summary of the dataset, avg number of constraints, cells, rows, columns}
In this section we experimentally validate our approach.
We studied various questions, most notably with what accuracy our algorithm can find constraints.

@@

There are three main categories of spreadsheets in the benchmark: spreadsheets from an exercise session teaching Excel, spreadsheets from tutorials online and publicly available spreadsheets such as crime statistics or financial reports that demonstrate more real world usage\footnote{(attached to to the submitted paper)}.

@@

\subsection{Benchmark}
There are three main categories of spreadsheets in the benchmark: spreadsheets from an exercise session teaching Excel at an affiliated University, spreadsheets from tutorials online and publicly available spreadsheets such as crime statistics or financial reports that demonstrate more real world usage\footnote{(attached to to the submitted paper)}.
The case study is also included.

All spreadsheets have been converted to CSV files.
Definitions of tables were added manually, providing a way to compare results and overcoming shortcomings in the table detection algorithm (in some harder cases groups where also provided manually).

\samuel{Revisit}
For every spreadsheet a ground-truth has been provided manually, specifying the essential (or original) constraints that are expected to be discovered.
We consider three types of constraints:
\begin{description}
  \item[Implemented] Constraint that have been implemented and are expected to be found (all constraints in Table~\ref{table:constraints})
  \item[Essential] Constraint that have or have not been implemented but could be found using our algorithm (e.g. fuzzy conditional sum)
  \item[Non-trivial] Constraints currently outside of the scope of this system (e.g. generic nested mathematical or logical formulas or n-ary constraints)
\end{description}

@@

\subsubsection*{Q1. How accurate is the approach?}
Fortunately, for the implemented constraint templates, our system achieves a recall of $1.0$, i.e. it is able to find all of the supported constraints (90) on the benchmark suite.
There are only three templates that have not been implemented yet, so w.r.t. the ground truth the total recall of our system is $0.97$.

>>>>

\begin{example}
Consider the (headerless) table~$T_1$ in Figure~\ref{fig:main_example}.
Its rows are not type consistent (i.e. they contain both numeric and textual data).
However, the table can be partitioned into five column-oriented blocks $B_1, B_2, B_3, B_4, B_5$, as shown in the figure ($B_1 = \range{T_1}{\rangeall}{1}$, $B_2 = \range{T_1}{\rangeall}{2}$, $B_3 = \range{T_1}{\rangeall}{\rangeto{3}{8}}$, \dots).
% {\small
% B_1 = \range{T_1}{\rangeall}{1},
% B_2 = \range{T_1}{\rangeall}{2},
% B_3 = \range{T_1}{\rangeall}{\rangeto{3}{8}},
% B_4 = \range{T_1}{\rangeall}{9},
% B_5 = \range{T_1}{\rangeall}{\rangeto{10}{11}}
\end{example}

@@

Instead, we divide this step into two parts: in the first part, we will not reason over individual (sub)blocks and their data, but rather over the properties of the maximal superblocks.
% and what combination of superblocks may lead to a valid argument assignment.\\
Consider Table 1 and 2 in Figure~\ref{fig:main_example} and the $\mathit{SUM}_{col}$ constraint template.
Instead of considering all possible subblocks of T1 and T2, we first reason over the properties of the superblocks.
For example, $B2$ and $B4$ are not numeric so they can be immediately discarded.
\samuel{Iff Br is from T2?}
Looking at the sizes, $B1$ and~$B2$ contain only one and two columns respectively, however, the rows in $T2$ have length 4 so we only need to consider superblocks with at least 4 columns, which only $B3$ satisfies.

In the second part, we start from an atom $c(B_1, \ldots ,B_n)$ with maximal superblocks $B_i$, and generate and test all possible (sub)block assignments to arguments using the properties and actual data of the blocks. As we only have to consider the blocks $B'_i \sqsubseteq B_i$ contained in each superblock, this is typically much easier. For the above example, each of the vectors of $B6$ will be considered as candidate for the left-hand side, and one can enumerate all subblocks for the right-hand side and verify the signature (e.g. at least size $4$) and test the definition for each of the rows. %In practice, this can be optimized further.
\\\\
\samuel{Section references not correct}
We now describe in more detail how the headerless tables are extracted and how the maximal superblocks are generated from them (step 1 and 2, Section 4.1), how the candidate superassignments are generated (step 3a, Section 4.2) and how the actual assignments are extracted from that (step 3b, Section 4.3). We then describe a number of optimizations that we perform to speed up step 3 and 4 (Section 4.4).

@@

Table~\ref{tbl:translation} shows how the conversion from requirements of the signature to CSP constraints happens:
Requirements on the lengths, orientations and tables of blocks can be directly enforced since they are invariant under block containment($\sqsubseteq$).
Typing requirements need to be relaxed to check only for base types (numeric and textual).
Minimum sizes can be directly enforced, but exact size requirements are relaxed to minimum sizes, since blocks with two many vectors contain subblocks of the required size.
Finally, restrictions on the number of rows or columns behave as length or size constraints based on the orientation of the block they are applied to.
Subgroups of row-oriented (column-oriented) blocks will always have the same number of columns (rows), however, the number of rows (columns) might decrease.

The $\generategroups(\textit{s,\blocks})$ method will use these conversion rules to construct a CSP and query a solver for all solutions.
These solutions correspond to the valid superblock assignments for constraint template~$s$.

@@

In this step, we could also formulate the problem as a CSP. %However, the search space is much smaller, e.g. $O(nm^2)$ with $n$ the number of arguments of the template and $m$ the size of the largest block.
However, few CSP solvers support floating point numbers, which are prevalent in spreadsheets.
Furthermore, in a CSP approach we would have to \textit{ground} out the definition for each of the corresponding elements in the vectors.
\samuel{Tias, there is something wrong here:}
This is inefficient, as the signature of the definition may already not be satisfied, or the first element in the data.

@@

\samuel{Angelika: trim this section for repeated information}
Using the dependency graph~(\dependencies) we can reorder the templates such that a template occurs after any template it depends on.
Concretely, the input templates~$\temps$ (see Algorithm~\ref{algo:tcl}) are reordered to agree with the partial ordering imposed by~\dependencies.

The dependency graph represents a partial ordering and is used to produce a linear ordering of the given templates such that more general templates occur before templates that depend on them (e.g. \textit{ALLDIFFERENT} before \textit{FOREIGNKEY}).
Considering Algorithm~\ref{algo:tcl}, this step can be seen as preprocessing the input~\temps.
Learning in this order allows templates to use constraints they depend on as these will have already been found.
This is exploited in the \generategroups step to speedup the algorithm by reducing the search space.
Let us illustrate this step with an example.

@@

\subsubsection{Redundancy}
We consider two types of redundancy that we aim to eliminate during the search.
As noted in Section~\ref{sec:form:dependencies}, for some constraint templates there are \textit{symmetries} over the arguments that lead to trivially equivalent solutions, for example, $\ecprod{B_R'}{B_1'}{B_2'} \Leftrightarrow \ecprod{B_R'}{B_2'}{B_1'}$.
Such duplicates are avoided by defining a \textit{canonical} form for these constraints.
In practice, we define an arbitrary but fixed block ordering and require those blocks that are interchangeable to adhere to this ordering.
Moreover, \sname currently contains only product (\ecprod{B_R'}{B_1'}{B_2'}) and not division (\eccalc{B_1'}{B_R' / B_2'}) as well as difference (\ecdiff{B_r}{B_1}{B_2}) and not sum (\eccalc{B_1}{B_r + B_2}).
These choices are based on the prevalence of product and difference in financial spreadsheet data, however, division and sum as well as other constraints could easily be added.

@@

\subsubsection{Constraints}
The constraints that are currently supported in our system are shown in Table~\ref{table:constraints}.
Their inclusion was based on the analysis of the most popular Excel constraints \samuel{How do we know?} as well as the spreadsheets we encountered.
All these constraints are available in popular spreadsheet software, except for the structural constraints (\textit{ALLDIFFERENT}, \textit{PERMUTATION} and \textit{FOREIGNKEY}).
These constraints were added so that they can be used in the signature of other constraint templates; they are also popular in constraint satisfaction~\cite{modelseeker}.

@@

\section{Evaluation}\label{sec:evaluation}
\sergey{we need to add a summary of the dataset, avg number of constraints, cells, rows, columns}
In this section we experimentally validate our approach.
We studied various questions \samuel{Mention questions here?}, most notably with what accuracy our algorithm can find constraints.

@@

There are three main categories of spreadsheets in the benchmark: spreadsheets from an exercise session teaching Excel \samuel{Including running example?}, spreadsheets from tutorials online and publicly available spreadsheets such as crime statistics or financial reports that demonstrate more real world usage\footnote{(attached to to the submitted paper)}.

@@

\subsection{Benchmark}
All spreadsheets have been converted to CSV files.
Definitions of tables were added manually, providing a way to compare results and overcoming shortcomings in the table detection algorithm (in some harder cases blocks where also provided manually).

\samuel{Revisit [what does "implemented" mean here? what's the difference between "implemented" and "algorithm", esp given that we are talking experiments? ]}
For every spreadsheet a ground-truth has been provided manually, specifying the essential (or original) constraints that are expected to be discovered.
We consider three types of constraints:
\begin{description}
  \item[Implemented] \samuel{Dupl} Constraint that have been implemented and are expected to be found (all constraints in Table~\ref{table:constraints})
  \item[Essential] \samuel{Dupl} Constraint that have or have not been implemented but could be found using our algorithm (e.g. fuzzy conditional sum)
  \item[Non-trivial] Constraints currently outside of the scope of this system (e.g. generic nested mathematical or logical formulas or n-ary constraints)
\end{description}

@@

\subsubsection*{Q1. How accurate is the approach?}
\samuel{Sounds like we were surprised}
Fortunately, for the implemented constraint templates, our system achieves a recall of $1.0$, i.e. it is able to find all of the supported constraints (90) on the benchmark suite.
There are only three templates that have not been implemented yet, so w.r.t. the ground truth the total recall of our system is $0.97$.
